---
title: An idea to visualize combination drug treatment experiments in R
author: Caglar Berkel
date: '2020-01-18'
slug: an-idea-to-visualize-combination-drug-treatment-experiments-in-r
categories:
  - Data science
tags:
  - Data Science
---

Recently, I have been thinking on how to visualize my experimental data where I performed combination drug treatments on certain cancer cell lines, for a paper I am writing. In short, I used two different inhibitor molecules alone or in combination, and measured a response variable (percent cell viability in my case). First, I drew side-by-side boxplots for each case (drugs alone in different concentrations and drugs in combination applied with previously selected concentrations) using ggplot and I thought it was not so easy for the reader to interpret the data as he/she should be. Then, while I was reading Hands-On Machine Learning with R by Bradley Boehmke & Brandon Greenwell, I got an idea on how to better visualize my data. I decided to use scatter3D function in plot3D R package. Here is an example of one of my figures and the code chunk to produce this 3D plot:

```{r}
library(readxl)
library(tidyverse)
library(plot3D)
```

```{r}
SRBdata <- read_excel("/Users/caglarberkel/Desktop/LAB/Experiment data/SRB data/CIS_BOR_OLA_all cells_SRB data.xlsx")
```

```{r}
df1 <- SRBdata %>% filter(CIS == 0) %>% filter(BOR == 0 | BOR == 10 | BOR == 20) %>% filter(OLA == 0 | OLA == 10 | OLA == 20) %>% filter(cellline == "OV2008")

x <- df1$BOR
y <- df1$OLA
z <- df1$viability
fit <- lm(z ~ x + y)

grid.lines = 26
x.pred <- seq(min(x), max(x), length.out = grid.lines)
y.pred <- seq(min(y), max(y), length.out = grid.lines)
xy <- expand.grid( x = x.pred, y = y.pred)
z.pred <- matrix(predict(fit, newdata = xy), nrow = grid.lines, ncol = grid.lines)
fitpoints <- predict(fit)
g1 <- scatter3D(x, y, z, pch = 18, cex = 2, 
           theta = 45, phi = 45, ticktype = "detailed",
           xlab = "BOR", ylab = "OLA", zlab = "% viability",  
           surf = list(x = x.pred, y = y.pred, z = z.pred,  
                       facets = NA, fit = fitpoints), main = "OV2008")
fit$coefficients
```



